### 2019-03-19 4_implementation.tex
More text on implementations.

For Task Bench, we decided to target a relatively low-level
implementation in Chapel, using explicit task instantiation (via
\lstinline[language=Chapel]{coforall}), bulk access to distributed
arrays for efficient data movement, and atomic integers for
synchronization. The advantage of this approach is that it permits
fine-tuning of the implementation for performance, and avoids relying
on advanced optimizations in the Chapel compiler for performance.

Chapel also supports a more high-level approach to parallel
programming in which the granularities of tasks are determined
automatically by the compiler (via
\lstinline[language=Chapel]{forall}). This style was not evaluated in
Task Bench as we were concerned that the dynamic, input-dependent
nature of the task dependencies would defeat optimizations in the
Chapel compiler that might turn out to be critical for performance.

@@

Our Task Bench implementation was straightforward: we create a chare
array for the task graph, with one chare for each column. Methods are
used to implement dependencies, and a task is executed as soon as its
dependecies are all available. Collectives are used to synchronize the
completion of the task graph execution.

@@

MPI~\cite{MPI}

@@

OmpSs~\cite{OmpSs11}

@@

OpenMP~\cite{OpenMPSpec40}

@@

PaRSEC~\cite{PARSEC13}

@@

Realm~\cite{Realm14}

@@

Regent~\cite{Regent15}

>>>>

For Task Bench, we decided to target a relatively low-level
implementation in Chapel, using explicit task instantiation (via
\lstinline[language=Chapel]{coforall}), bulk access to distributed
arrays for efficient data movement, and atomic integers for
synchronization. The advantage of this approach is that it permits
fine-tuning of the implementation for performance, and avoids relying
on advanced optimizations in the Chapel compiler, which might struggle
with the dynamic, input-dependent nature of the task graphs in Task
Bench.

Chapel also supports a more high-level approach to parallel
programming in which the granularities of tasks are determined
automatically by the compiler (via
\lstinline[language=Chapel]{forall}). This is the level that users
would typically target in their programs. The specific dependence
patterns supported by Task Bench might also be implementable at this
level, though a full Task Bench implementation would be challenging.

% FIXME: Does the Chapel compiler handle: bulk data movement, point-to-point synchronization

Internally, Chapel implements custom support for Cray networks, and
uses the GASNet~\cite{GASNET07} compatibility layer on other networks.

@@

Our Task Bench implementation is straightforward: we create a chare
array for the task graph, with one chare for each column. Methods are
used to implement dependencies, and a task is executed as soon as its
dependecies are all available. Collectives are used to synchronize the
completion of the task graph execution.

% FIXME: Discuss SMP vs non-SMP versions

Internally, Charm++ implements support for a variety of networks
directly on top of the low-level, vendor-provided APIs. For Cori, we
use Charm++'s native support for uGNI.

@@

MPI~\cite{MPI} is the industry standard message-passing API for
programming supercomputers, and is the most widely used programming
model in high-performance scientific computing today. The goal of our
MPI implementation of Task Bench is to represent the common (if
unexciting) case which continues to dominate high-performance
computing: bulk synchronous programming with distinct computation and
communication phases. We use \lstinline[language=C++]{MPI_Isend} and
\lstinline[language=C++]{MPI_Irecv} to implement the communication
phase, posting receives before sends. Each task dependency maps to one
send/receive pair in MPI. Despite the relatively straightforward
design, our MPI implementation is among the best performing.

In our evaluation, we use the default MPI implementation installed on
the machine, which is Cray MPI on the Cori supercomputer. This
implementation is vetted by the machine operators and vendor and is
the highest performing available implementation for the hardware.

For completeness, we also explored the use of additional MPI calls:
namely buffered sends (\lstinline[language=C++]{MPI_Bsend}),
broadcasts (\lstinline[language=C++]{MPI_Bcast}), and all-to-all
communication (\lstinline[language=C++]{MPI_Alltoallv}). For the
configurations used in our evaluation, we found non-blocking sends and
receives to be the best performing, often by a wide margin.

@@

OmpSs~\cite{OmpSs11} is a programming model for loop-based parallelism
that is source-compatible with OpenMP. Our Task Bench implementation
uses OpenMP 4.0-style tasks with task dependencies. Because tasks in
OpenMP 4.0 have a fixed number of dependencies, we use a switch
statement to implement the dynamic dependencies required for Task
Bench. The implementation is otherwise straightforward.

% FIXME: More to say about OmpSs? Is there anything different vs the OpenMP implementation?

@@

OpenMP~\cite{OpenMPSpec40} is the industry standard API for loop-based
parallelism on shared-memory systems. Our Task Bench implementation
for OpenMP is very similar to OmpSs and uses OpenMP 4.0 tasks with
task dependencies.

In our evaluation, we test with both GNU GOMP and Intel KMP
implementations of OpenMP. We found KMP to provide better performance
and thus report KMP numbers in our evaluation.

@@

PaRSEC~\cite{PARSEC13} is a task-based programming system which
supports two distinct programming models: \emph{parameterized task
  graphs} (PTG) and \emph{sequential task flow} (STF). In the PTG
model, the program consists of an analytical function which describes
the expansion of the task graph. PTG is therefore a more limited
programming model, but can potentially provide better performance
because the graph description is amenable to static analysis and can
be pruned efficiently for horizontally scalable execution.

In the STF model, tasks are enumerated in program order (by
dynamically executing the program), and dependencies between tasks are
identified automatically based on the arguments to tasks. A task
depends on another task if it reads data written by the other task,
and the data is copied automatically if the two tasks are executed on
different nodes. While the basic STF model is implicitly parallel
because the system is responsible for the discovery of parallelism, in
order to achieve horizontal scalability, the program is executed in a
SPMD fashion on all nodes, and the user is responsible for eliding
tasks that are not directly connected to those that are to be executed
on the current node.

Data in PaRSEC is stored in \emph{regions}, or dense arrays of
contiguous memory. Regions may be passed to tasks on other nodes, but
can only be written on the node where the data is originally
allocated. In our Task Bench implementation, each task writes to a
region as output, and these regions are passed as inputs to the
subsequent set of tasks. As in the OmpSs and OpenMP implementations,
the task launch code uses a switch statement as the PaRSEC API assumes
tasks have a fixed number of dependencies.

% FIXME: What does PaRSEC call regions?

Internally, PaRSEC uses MPI for network communication.

@@

Realm~\cite{Realm14} is an explicitly-parallel task-based programming
model that is used internally by Legion~\cite{Legion12} and
Regent~\cite{Regent15}, which are implicitly parallel layers written
on top of Realm. Thus the Realm implementation of Task Bench can be
seen as a limit study of what is possible with Legion or Regent.

Tasks in Realm are executed in a deferred manner, with dependencies
between tasks explicitly defined by \emph{events} passed from one task
to another. Realm's data model supports regions which live on a single
node. Data must be explicitly copied, and the copies synchronized with
tasks via events.

Realm also supports a subgraph API where sets of tasks and copies are
issued together. This mode permits additional optimizations to be
performed. We have implemented Task Bench in both individual and
subgraph styles, but report numbers for the subgraph version as it
provides better performance.

Internally, Realm uses GASNet~\cite{GASNET07} for communication.

@@

Regent~\cite{Regent15} is an implicitly-parallel task-based language
which implements the Legion programming model~\cite{Legion12}. We use
Regent rather than Legion directly because the Regent compiler
provides a critical optimization, \emph{control replication}, that
enables superior horizontal scalability of
applications~\cite{ControlReplication17}.

In Regent tasks are executed sequentially and parallelism is
discovered automatically based on the data that is read or written by
tasks. Regent programs are agnostic to the machine architecture and
are separately \emph{mapped} for execution on a distributed memory
system. Copies are inserted automatically by the compiler and runtime
whenever data is read on a different node from where it is written.

The Regent compiler performs control replication to slice the Regent
program into \emph{shards} of execution which can scalably execute on
large numbers of nodes. Our Regent implementation also makes use of a
tracing optimization~\cite{LegionTracing18} which substantially
reduces the cost of dynamic analysis of tasks in Regent.
